Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-18 special-issue-ci/Response to reviewers.tex

You are absolutely right. We have corrected that paragraph to include
the actual contents of the paper.

\begin{quote}
Summarizing the review, the authors should provide a convincing example showing the advantages of
their proposal. As presented, the paper is a proof of concept of a system that seems difficult to be used
by the researchers needing high computational resources to run an evolutionary algorithm to solve a realworld
problem.
\end{quote}

We have added a new experiment (Section V.A), a more complex and
taxing problem that also uses a web-worker architecture, in order to
prove how this can be 
used in real-life problems. In this case, we have achieved many more
simultaneous users, up to 25 in some minutes and more than 100 in a
single hour. All in all, more than 400 volunteers (identified by
unique IPs) have participated in
the experiment.

In fact, our architecture {\sf NodIO} is
not an exclusively browser-based one. Its pool-based architecture allows
any other kind of client, including local ones, to be incorporated to
the experiment, with the lightweight overhead of doing asynchronous
HTTP requests, which being local are indeed negligible.

### 2015-11-18 special-issue-ci/Response to reviewers.tex

\noindent
Yours sincerely,\\
The authors.

\section{Response to Reviewer \#1}

\begin{quote}
The main contribution of this paper is the proposal of a Javascript based distributed system intended to
run evolutionary algorithms according to the volunteer computing scheme. The idea is that user can
contribute to the computation by using only a browser, without needing to install any plugin nor a Java
virtual machine.
The volunteer computing model appeared some years ago and it has been used in many fields. The
model is effective to solve problems demanding high computational capabilities, and at this point is where
this paper proposal introduces some concerns. The choice of Javascript code running on browsers has
some advantages, as is pointed out in the paper, but some important drawbacks. Given that the
motivation of these kinds of distributed models is to solve complex problems, using Javascript doesn\'{}t
appear a right choice. It\'{}s true that is a very commonly used programming language, but in the context of
Web applications, not in high performance ones. Apart from the fact of being an interpreted language
designed for the Web, it is difficult to think on a computer scientist writing a complex math code in
Javascrip. If the authors would have used as a case study a complex optimization problem then they
could have provided numerical data to sustain their claims; however, the selected problem (which should
be formulated in the paper) seems a rather academic one.
\end{quote}

We thank the reviewer for the comment. The only reason why JavaScript
was chosen was because it is the only language that is native to the
browser, allowing to carry out precisely this kind of experiments
where the user is invited to click on an URL and start giving CPU to
it. Although there are many other volunteer frameworks, written in
other languages, that a priori would offer a higher performance, the
key is that they would {\em while they are running}, that is, as soon
as it has been installed, the user has decided to fire it up, and it
does not affect the performance of the rest of his system, which is a % se confundia con el desempeño del todo el sistema
concern in this kind of systems. We have performed speed tests on
JavaScript and other compiled languages resulting that, in fact,
JavaScript uses a JIT compiler that provides a good performance; we
have published these tests as an open access technical report. \cite{2015arXiv151101088M}

@@

In a nutshell, what we are saying is that, as is proven in the paper,
a JavaScript application, running in the browser, can gather in
seconds dozens of users. That would be, in practice, impossible with
other kind of native clients, who would have to download and maybe
compile the application, losing many seconds or minutes for the
optimization problem that has to be solved.

These facts, along the explanations, have now been added to the first
paragraph of the {\em Description of the system} section, as well as
to the introduction.

\begin{quote}
The paper organization is not clear. Some comments about this:
- The title "Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed
Evolutionary Computation" seems to suggest that the volunteers run in the cloud, but that is not the case:
the server run in the cloud and the volunteers do in the browsers of
common PCs and laptops.
\end{quote}

That is correct. We just used the term {\em cloud} in a relatively informal
sense because they are available as in a cloud for running the
experiment. That is, from the point of view of the person running the
experiment, there is a {\em cloud} out there running your experiments,
which is actually run by volunteers in their browsers. We went to the
original acceptation of ``cloud'' as a ``network'' more than the more
precise and technical acceptation used today. %PABLO: no existe aception, lo he cambiado a acceptation (segun WordReference)

@@

\end{quote}

We have inserted a paragraph at the end of the section that describes
the experimental setup describing what are the challenges that this
kind of voluntary framework must overcome. Another paragraph analyzing
the security from the point of view of the framework used has been
also added. A new paragraph on the security challenges has also been
inserted, third paragraph in this section, to clarify our integral and
pragmatic approach to security, which so far has worked perfectly.

\begin{quote}
-  an experiment with a more difficult problem (please use the well-known benchmark problems
described in literature).
\end{quote}

An experiment based on the the well known Rastrigin's function has been
performed and analyzed. It is now Subsection V.A.

@@

\begin{quote}
-  A comparison in similar conditions of the two implementations of the architecture (the algorithms
used in the experiments are slightly different, but this could lead to dramatic changes in
performance, and consequently, invalidate the comparison performed in the paper). This would
enable a measurement of the actual improvement in performance obtained by the use of web
workers (not that generated due to a different algorithm)
\end{quote}

This has been included as the last subsection of the paper previous to
the conclusions. The new
experiment has a constant population instead of using  population of random size.
We have not found a big influence on the result.

@@

JavaScript is not, in fact, so slow and to prove its speed we have
published a technical report comparing it with other compiled
languages (such as Scala) and scripted languages
\cite{2015arXiv151101088M}. In fact, node.js (used for the benchmarks)
and the JavaScript virtual machine in Chrome (used in Android phones
and Chrome and Chromium browsers) share the same JIT compiler, so
their speeds are virtually the same. In general, as a conclusion of
the report we see that the real difference between scripted and
compiled languages will vary with the language, but also with the
particular implementation of the algorithm and the size of the problem
or chromosome you are using. Even if the speed for a particular
problem is two or five times that achieved for JavaScript, this
disadvantage is more than offset by the availability of that free
resource in abundance; that is, you will only need that many
volunteers cooperating to get better performance than what you would
get with C or Java. Even so, this NodIO architecture, as highlighted
in the new version of the paper, is pool-based so local sources using
high-speed devices such as GPUs or even local powerful CPUs can be
added as non-volunteer users, since the heterogeinity and simple API
of the system allows that by design.

\begin{quote}
The paper is well written and organized. State of the art is quite complete. Nevertheless, the experimental
section must be carefully checked and extended in order to fully validate the goodness of the proposed
architecture. At least, authors should show the behavior of the architecture in a real-world problem
demanding high computational performance.
Finally, English must be double-checked in order to correct minor typos such as "in the sense than[that]"
(in page 2).
\end{quote}

We have extended the experiments section with a new one, based on a
hard version of the Rastrigin's problem.

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adding meself as author, adding @mariosky explanation about the feature of first NodIO version

\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$
\thanks{Manuscript submitted for review on \today.}%
\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
\thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es}}%
\thanks{*Corresponding author.}%
}

@@

A summary of the results of each run is also shown in Table
\ref{tab:summary:os}, which shows the median number of Internet Protocol addressess (IPs)
intervening in each experiment,  median time needed
to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
then two columns showing the percentage of experiments that took less
than the two baseline experiments shown in the introduction to this
section. The first striking result is that in all cases, 50\% of the
experiments involved 5 or less IPs. This is consistent with results
previously obtained \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
the usual number of IPs that participated in an experiment. The
maximum number of different IPs for each experiment is also in the
same range and of the order of 10, which is also consistent with
prior work. The median time has a big range of variation, but 50\% of
the time takes less than several minutes, from around 4 minutes in the
best case to roughly 2/3 of an hour in the worst case. That figure is
not competitive, {\em a priori}, with the baseline experiment, that is
why it is interesting to look at the columns which measure
the percentage of experiments in which the solution was found in less
time than the baseline. It is similar in all three cases, with around
20\% on average for the first baseline experiment, roughly 3\% in the
second case. The comparison is not totally fair: this experiment takes
place in a browser, and in fact most of the time is devoted to
presenting the charts; the server is also not the same since it is run in a
PaaS instead of a high-performance desktop computer, and it is also a
remote server, not a local server as in the baseline experiment. But
even in this case, we were looking for creating a framework that
allowed us to run evolutionary computation experiments {\em faster} than
in our very own computer. And the results indicate that although it will
happen in 1 out of roughly 5 or roughly 20 cases, depending on what baseline
experiment you choose, it will not happen always or even most of
the time. One of the problems is shown in the last column, which
indicates the statistical correlation between the number of IPs
participating in one experiment and the next. In all cases,
correlation is so low as to affirm that there is almost no
relationship between them, that is, computing nodes are not {\em
  staying} after the solution has been found. This is a feature of
this implementation: the user has to actively reload the page to make
the evolutionary algorithm start again. % (Paloma) Feature or disadvantage?
 However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
unnecessary upper limit on the time, or number of experiments, that a user
will voluntarily perform.

@@

The experiments were run in the same way as before, using several
Twitter announcements throughout several days by the end of July
2015. Eventually, more than one thousand experiments were completed in
a matter of days. A summary of results is shown in Table
\ref{tab:summary:ww}, comparing it with the aggregate of the results
obtained with the initial version of {\sf NodIO}. These results are
remarkably different, being similar only in the median and maximum
number of IPs, although in this case it would be combination
IP-worker, as we considered every worker as a unique IP. The median
number of seconds is almost 80\% better than the previous version, and
the median number of PUTs is three times better; this causes that most
of experiments take less time than one of the baselines, and one third
less than the second baseline. Thus, the conclusion in this case is
that we can obtain, through volunteer computation, a result that is,
most of the times, better than we would using a similar, non-parallel,
desktop setup, which makes this system suitable for massively
distributed evolutionary algorithms.

@@

This is in part due to the fact that the presence of nodes is not
constant, as shown in Figure \cite{fig:ips:rastrigin}, which plots the
number of unique IPs participating in the experiment per minute (left)
and hour (right). The two graphs have the same shape which matches
also the speed in which fitness is decreased in Figure
\ref{fig:puts:rastrigin}. At the beginning there are very few or only
one, but later one there ar peaks of almost 30 unique machine a
minute; these peaks appear again when new announcements are made in
Twitter or any other social networks, to go down to a few remaining
nodes later on. It is interesting to note that, at the most, more than
100 unique nodes were participating in the experiment during one hour;
also that there are sustained peaks during a day of 6-7 computers
every minute. This is due to the fact that the computation is
asynchronous, but also to the fact that different computers stay in
the simulation for a different period of time. This period has been
plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
{\tt PUT}s per node, ranked by number, along with a fit to the two
distributions we have used before, GEV and Weibull; the parameters of
this fit are shown in Table \ref{tab:puts:ww:f15}. Unlike the previous
experiments, in which GEV fit was relatively good, in this case it is
quite obviously not. However, the fit to the Weibull distribution,
shown at the left of the image, is relatively good. As should be
expected, the shape and scale are different to the one in the previous
experiments with the Trap function, however, the shape is quite
similar (0.40 vs 0.69), being in both cases less than one, which
indicates a decreasing distribution of the same kind. In fact, the
difference in the scale parameter indicates that the difference
between the time devoted by a volunteer and the next is bigger in this
case. This is probably due to the fact that the duration of the
experiment is bigger, allowing for bigger differences in
behavior. While the first in the Trap function might have devoted a
few minutes and the second a few seconds less, the difference between
the first and second in this experiment might be in the scale of
hours.

>>>>

\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$
\thanks{Manuscript submitted for review on \today.}%
\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
\thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es}}%
\thanks{*Corresponding author.}%
}

@@

A summary of the results of each run is also shown in Table
\ref{tab:summary:os}, which shows the median number of Internet Protocol addressess (IPs)
intervening in each experiment,  median time needed
to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
then two columns showing the percentage of experiments that took less
than the two baseline experiments shown in the introduction to this
section. The first striking result is that in all cases, 50\% of the
experiments involved 5 or less IPs. This is consistent with results
previously obtained \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
the usual number of IPs that participated in an experiment. The
maximum number of different IPs for each experiment is also in the
same range and of the order of 10, which is also consistent with
prior work. The median time has a big range of variation, but 50\% of
the time takes less than several minutes, from around 4 minutes in the
best case to roughly 2/3 of an hour in the worst case. That figure is
not competitive, {\em a priori}, with the baseline experiment, that is
why it is interesting to look at the columns which measure
the percentage of experiments in which the solution was found in less
time than the baseline. It is similar in all three cases, with around
20\% on average for the first baseline experiment, roughly 3\% in the
second case. The comparison is not totally fair: this experiment takes
place in a browser, and in fact most of the time is devoted to
presenting the charts; the server is also not the same since it is run in a
PaaS instead of a high-performance desktop computer, and it is also a
remote server, not a local server as in the baseline experiment. But
even in this case, we were looking for creating a framework that
allowed us to run evolutionary computation experiments {\em faster} than
in our very own computer. And the results indicate that although it will
happen in 1 out of roughly 5 or roughly 20 cases, depending on what baseline
experiment you choose, it will not happen always or even most of
the time. One of the problems is shown in the last column, which
indicates the statistical correlation between the number of IPs
participating in one experiment and the next. In all cases,
correlation is so low as to affirm that there is almost no
relationship between them, that is, computing nodes are not {\em
  staying} after the solution has been found. This is a feature of
this implementation: the user has to actively reload the page to make
the evolutionary algorithm start again. Thus, the user can decide if to continue sharing resources or not.% (Paloma) Feature or disadvantage?
 However, it puts a probably            % (Mario) Feature if you only want to share your CPU a little  
unnecessary upper limit on the time, or number of experiments, that a user
will voluntarily perform.

@@

The experiments were run in the same way as before, using several
Twitter announcements throughout several days by the end of July
2015. Eventually, more than one thousand experiments were completed in
a matter of days. A summary of results is shown in Table
\ref{tab:summary:ww}, comparing it with the aggregate of the results
obtained with the initial version of {\sf NodIO}. These results are
remarkably different, being similar only in the median and maximum
number of IPs, although in this case it would be combination
IP-worker, as we considered every worker as a unique IP. The median
number of seconds is almost 80\% better than the previous version, and
the median number of PUTs is three times better; this causes that most
of experiments take less time than one of the baselines, and one third
less than the second baseline. Thus, the conclusion in this case is
that we can obtain, through volunteer computation, a result that most of the times is better than we would by using a similar, non-parallel,
desktop setup, which makes this system suitable for massively
distributed evolutionary algorithms.

@@

This is in part due to the fact that the presence of nodes is not
constant, as shown in Figure \cite{fig:ips:rastrigin}, which plots the
number of unique IPs participating in the experiment per minute (left)
and hour (right). The two graphs have the same shape which matches
also the speed in which fitness is decreased in Figure
\ref{fig:puts:rastrigin}. At the beginning there are very few or only
one, but later one there ar peaks of almost 30 unique machine a
minute; these peaks appear again when new announcements are made in
Twitter or any other social networks, to go down to a few remaining
nodes later on. It is interesting to note that, at the most, more than
100 unique nodes were participating in the experiment during one hour;
also that there are sustained peaks during a day of 6-7 computers
every minute. This is due to the fact that the computation is
asynchronous, but also to the fact that different computers stay in
the simulation for a different period of time. This period has been
plotted in Figure \ref{fig:fit:rastrigin}, that plots the number of
{\tt PUT}s per node, ranked by number, along with a fit to the two
distributions we have used before, GEV and Weibull; the parameters of
this fit are shown in Table \ref{tab:puts:ww:f15}. Unlike the previous
experiments, in which GEV fit was relatively good, in this case it is
quite obviously not. However, the fit to the Weibull distribution,
shown at the left of the image, is relatively good. As should be
expected, the shape and scale are different to the one in the previous
experiments with the Trap function, however, the shape is quite
similar (0.40 vs 0.69), being in both cases less than one, which
indicates a decreasing distribution of the same kind.
